---
title: "Recommender 3"
author: "Yadu"
date: "February 22, 2017"
output: html_document
---

## Motivation

The purpose here is to implement singular value decomposition to build a content-based recommender system.

## Data Creation

We will build a toy dataset as follows. 

```{r}
moviesraw <- matrix(c(4, NA, 3, 5, 2, NA, 5, NA, 5, 4, NA, 4, 5, 3, 5, 4, 2, NA, 1, 1, 4, 2, 4, NA, 3, 5, 3, 5, 3, 4, 5, NA, NA, 5, NA, NA, 1, 3, 5, 5, NA, 4), nrow = 6, byrow = T)
colnames(moviesraw) <- c("Batman Begins", "Alice in Wonderland", "Dumb and Dumber", "Equilibrium", "Matilda", "Salt", "Marley and Me")
rownames(moviesraw) <- c("Adam", "Benjamin", "Charlie", "David", "Edward", "Matthew")
moviesraw
```

## Checking the Spread

First, we will obtain the similarity matrix and perform singular value decomposition to determine underlying relationships. 

```{r}
library(recommenderlab)
m <- as(moviesraw, "realRatingMatrix")
simm <- as.matrix(similarity(m, method = "cosine", which = "items"))
simm
```

```{r}
s <- svd(simm)
s$u #left singular matrix
```

```{r}
V <- s$v
t(V) #right singular matrix
```

Due to the great variation in the eigenvalues, it can be inferred that dimension reduction is possible.

```{r}
s$d #eigenvalues
```

## Creating a Content-Based Recommender

First, we will build a toy vector that contains text information about these movies hardcoded from imdb. 

```{r}
reviews <- c("After training with his mentor, Batman begins his fight to free crime-ridden Gotham City from the corruption that Scarecrow and the League of Shadows have cast upon it.", "Nineteen-year-old Alice returns to the magical world from her childhood adventure, where she reunites with her old friends and learns of her true destiny: to end the Red Queen's reign of terror.", "Twenty years since their first adventure, Lloyd and Harry go on a road trip to find Harry's newly discovered daughter, who was given up for adoption.", "In a fascist future where all forms of feeling are illegal, a man in charge of enforcing the law rises to overthrow the system.", "Story of a wonderful little girl, who happens to be a genius, and her wonderful teacher vs. the worst parents ever and the worst school principal imaginable.", "A CIA agent goes on the run after a defector accuses her of being a Russian spy.", "A family learns important life lessons from their adorable, but naughty and neurotic dog.")
reviews
```

Next, the `tm` package is used for text-mining where a corpus for each movie is created. 

```{r}
library(tm)
movie_corpus <- Corpus(VectorSource(reviews))
```

Now a document-term matrix is created to determine the importance of each movie. Then, TF-IDF weighting is used adjusting for the fact that certain words appear more often.

```{r}
dtm <- DocumentTermMatrix(movie_corpus, control = list(removePunctuation = TRUE, removeNumbers = TRUE, stopwords = TRUE, tolower = TRUE, weighting = weightTfIdf))
dtm_matrix <- as.matrix(dtm)
dimnames(dtm_matrix)$Docs <- colnames(moviesraw)
```

## Performing Singular Value Decomposition on Content

Singular value decomposition is performed as follows. 

```{r}
movie_svd <- svd(dtm_matrix)
Sigmak <- movie_svd$d; Uk <- movie_svd$u; Vk <- t(as.matrix(movie_svd$v))
```

## Creating the System

The reduced transposed matrix below is the basis for this recommender system with the columns representing the 5 movies and the rows representing the 5 singular values. 

```{r}
movie_cat <- as.matrix(t(Uk))
dimnames(movie_cat) <- list(SVs = paste0("sv", 1:7), Movies = colnames(moviesraw))
```

A similar matrix of users and categories can be created now after normalizing the toy dataset and then creating a mean-centered matrix. 

```{r}
library(recommenderlab)
movies <- as(moviesraw, "realRatingMatrix")
movies_norm <- normalize(movies)
norm_cat <- as.vector(movies_norm@data)
norm_cat <- ifelse(norm_cat == 0, NA, norm_cat)
norm_cat <- matrix(norm_cat, nrow = nrow(movies), ncol = ncol(movies))

user_cat <- matrix(nrow = nrow(moviesraw), ncol = ncol(moviesraw))
for (i in 1:nrow(user_cat)){
  for (j in 1:ncol(user_cat)){
    user_cat[i,j] <- sum(norm_cat[i,]*movie_cat[j,], na.rm = T)/sum(movie_cat[j,] != 0) 
  }
}
```

Now that both matrices have been created, Pearson's correlation can be used to estimate the ratings for unrated movies.

```{r}
estimated <- matrix(nrow = nrow(moviesraw), ncol = ncol(moviesraw))
colnames(estimated) = colnames(moviesraw)
rownames(estimated) = rownames(moviesraw)
for (i in 1:nrow(estimated)){
  for(j in 1:ncol(estimated)){
    if(is.na(norm_cat[i,j])){
      for(k in 1:nrow(movie_cat)){
        num <- sum(user_cat[i,k]*movie_cat[k,j])
        den <- sqrt(sum(user_cat[i,]^2, na.rm = T))*sqrt(sum(movie_cat[,j]^2, na.rm = T))
      }
      estimated[i,j] <- num/den
    }else{estimated[i,j] <- NA}
  }
}
estimated
```

SVD is not a suitable method to implement when building a content-based recommender system. The reason for this is because the ratings for this matrix are supposed to be between 1 and 5. There are ratings that are 0 as well as ratings that are close to 0. It seems that content-based recommendations would not suffice. 